---
title: "R Notebook"
output: github_document
---

```{r}
# libraries
library(tidyverse)
library(palmerpenguins)
library(ggplot2)
```

Problem 1
```{r}
set.seed(123)
df <- tibble(
  x_1 = rnorm(10, sd = 1),
  x_2 = x_1 > 0,
  x_3 = c("a","b","c","d","e","f","g","h","i","j"),
  x_4 = factor(c("1","2","3","2","3","1","1","3","2","1"))
)
mean_num = mean(pull(df,x_1))
mean_log = mean(pull(df,x_2))
mean_char = mean(pull(df, x_3))
mean_fact = mean(pull(df, x_4))
```
The means are shown as: `r mean_num`; `r mean_log`; `r mean_char`; `r mean_fact`. \
Mean of **numeric and logical vectors** can be calculated while mean of **character vector and factor vector** cannot be calculated. And thus, we got NAs for these two columns.
```{r}
as.numeric(df$x_2)
as.numeric(df$x_3)
as.numeric(df$x_4)
```
Logical vectors can be converted to numeric form with values 0 and 1; character values cannot be converted to numeric forms and thus return NAs; factor values can be converted to numeric forms according to their levels. \

Problem 2
```{r}
data("penguins", package = "palmerpenguins")
summary(penguins)
n_row = nrow(penguins)
n_col = ncol(penguins)
# strip NA values before calculating mean
m_flipper = mean(penguins$flipper_length_mm, na.rm = TRUE)
```
The penguins dataset contains **`r n_row` rows and `r n_col` columns**. The variables are: `r colnames(penguins)`. There are overall **3 species on 3 islands**, summary statistics are displayed in the table above.\
The mean of flipper length in mm is **`r m_flipper`**.

```{r}
plot_df <- tibble(
  flipper_length = penguins$flipper_length_mm,
  bill_length = penguins$bill_length_mm,
  fact = factor(penguins$species)
)
```
```{r}
plot <- ggplot(plot_df, aes(x = flipper_length, y = bill_length, color = fact)) + geom_point() + labs(title = "flipper length vs bill length") + theme(plot.title = element_text(face = "bold", hjust = 0.5))
print(plot)
ggsave("hw1_scatterplot.png", width = 5, height = 5)
```


